---
title: "HW 01"
author: "Yashi Mi"
format: 
  html:
    embed-resources: true
execute: 
  warning: false
  message: false
  error: false
toc: true
---

## 0 - Setup

```{r setup}
if (!require("pacman")) 
  install.packages("pacman")

# use this line for installing/loading
pacman::p_load (tidyverse,
              palmerpenguins )

devtools::install_github("tidyverse/dsbox")

library(here)
```

## 1 - Road traffic accidents in Edinburgh

```{r label-me-1}
accidents <- read_csv(here("data","accidents.csv"))
glimpse(accidents)
```

```{r}
accidents <- accidents |>
  mutate(day_type = case_when(
    day_of_week %in% c("Saturday", "Sunday") ~ "Weekend",
    day_of_week %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "Weekday",
    TRUE ~ as.character(day_of_week)))
ggplot(accidents, aes(x = time, fill = severity)) +
  geom_density(alpha=0.6) +
  facet_wrap(~day_type, ncol=1) +
  scale_fill_manual("Severity", values = c("#8E6C8A", "#5E9CA0", "#F4E04D")) +
  labs(
    title = "Number of accidents throughout the day",
    subtitle = "By day of week and severity",
    x = "Time of day",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)
```

::: callout-note
The figure illustrates the distribution of road traffic accidents in Edinburgh across different times of day, separated by weekday and weekend, and categorized by severity (Fatal, Serious, Slight).

On weekdays, the number of serious and slight accidents peaks between 16:00 and 18:00, likely corresponding to evening rush hours. Fatal accidents show a smaller peak around midday (12:00).

On weekends, the number of slight accidents also peaks around 16:00, but serious accidents tend to peak slightly later, between 18:00 and 20:00. Notably, there are no fatal accidents reported on weekends in this dataset.
:::

## 2 - NYC marathon winners

```{r label-me-2}

```

## 3 - US counties

## 4 - Rental apartments in SF

## 5 - Napoleon’s march.
